Representing Scenes as Light Fields for Real Time Rendering in Virtual Reality Display

### 2021-01-28 TOG/0-abstract.tex
last day

Traditional high-quality 3D graphics requires large volume of detailed data to be rendering. This compromises efficiency and local storage. In recent years, cloud-based streaming and neural representations have significantly revolutionized the pipeline by redistributing and compressing the computation/storage. However, they typically suffer from high latency and low quality for practically large virtual scenes with 6 degree-of-freedom camera motions. These problems are particularly exacerbated in immersive medias such as virtual and augmented reality due to the demand of high frequency and resolution.

Tailored for the future cloud-based, portable, and low-storage VR platforms, we present the first gaze-contingent 3D neural representation and view synthesis approach. Our method mutually bridges both visual acuity and neural representation for 6DoF immersive interaction. Moreoever, it optimizes between temporal computational latency and spatial predictive precision. Both objective analysis and subjective studies evidenced our effectiveness of significantly reducing local storage (\warning{xxx MB, xx\% size reduction} for large 3D scenes) while preserves original perceptual quality without introducing additional latency.
\end{abstract}

>>>>

Traditional high-quality 3D graphics requires large volumes of fine-detailed scene data for rendering. This demand compromises computational efficiency and local storage resources. Specifically, it becomes more concerning for future wearable and portable virtual and augmented reality (VR/AR) displays. Recent approaches to combat this problem include remote rendering/streaming and neural representations of 3D assets. These approaches have redefined the traditional local storage-rendering pipeline by distributed computing or compression of large data. However, these methods typically suffer from high latency or low quality for practical visualization of large immersive virtual scenes, notably with extra high resolution and refresh rate requirements for VR applications such as gaming and design.

Tailored for the future portable, low-storage, and energy-efficient VR platforms, we present the first gaze-contingent 3D neural representation and view synthesis method. We incorporate the human psychophysics of visual- and stereo-acuity into egocentric neural representation of 3D scenery. Furthermore, we jointly optimize the latency/performance and visual quality, while mutually bridging human perception and neural scene synthesis, to achieve perceptually high quality immersive interaction. Both objective analysis and subjective study demonstrate the effectiveness of our approach in significantly reducing local storage volume and synthesis latency (up to 99\% reduction in both data size and computational time), while simultaneously presenting high-fidelity rendering, with perceptual quality identical to that of fully locally stored and rendered high-quality imagery.
\end{abstract}

### 2021-01-28 TOG/1-introduction.tex
last day

\note{
What problem we are trying to solve.
Why it is important, and why people should care.
}%note

Started from remote desktops and online videos, cloud-based computing has show unprecedented potential to reshape computer graphics pipeline by completely relieving end-user's computational workload thus device bulkiness. In recent years, we have observed growing interactive streaming services such as Google Stadia or NVIDIA Shield. 
However, these off-the-shelf services have only been deployed in less latency sensitive platforms such as TV-based game consoles, due to the inevitable latency (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}. Immersive medias, such as virtual and augmented realities, demand remarkably lower latency and higher fidelity/speed for sickness-free and natural experience. Due to the consequently massive spatial-temporal data volume, there is no solution for streaming frames from server to individual end users today.

\note{
What prior works have done, and why they are not adequate.
(Note: this is just high level big ideas. Details should go to a previous work section.)
}%note

\note{
What our method has to offer, sales pitch for concrete benefits, not technical details.
Imagine we are doing a TV advertisement here.
}%note

This document provides a template for writing research papers, and describes how to manage project progress via iterating drafts.
Doing this correctly can help you achieve productive research and happy life, e.g. working anywhere anytime without synchronous meetings with your collaborators.

@@

I learned from my PhD adviser to start writing from the moment I have the faintest idea, and gradually update the drafts to reflect progress.
During my first job with NVIDIA, I learned why and how revision control is essential for sharing and editing source files, especially among large engineering teams.
The iterative nature of writing and coding matches well with the capability of revision control.
Thus I prefer Latex + git/svn (e.g. github, bitbucket, and the svn server under my hosting service), even though I can still manage with MS Word + cloud drive (e.g. Google drive, Box, and Dropbox). 
A good computer scientist writes papers like programs, and manages Latex files like source codes.
These repos are external memories and communication mediums for the collective brains of my teams.

@@

\input{example_fig}

Translating thoughts into words, diagrams, equations, pseudo-codes, or codes is a good way to clarify and consolidate.
If you are writing a graphics/HCI paper, design the figures so that they alone can provide a high level picture of the main points.
As exemplified in \Cref{fig:example}, a figure can contain images, drawings, or combinations.

@@

You can use the associated files as a starting template for your research papers.
Start with the \filename{Makefile}, and notice the two build targets: final for official submission and public disclosure, and draft for internal sharing among your collaborators (including yourself).

>>>>

\note{
What problem we are trying to solve.
Why it is important, and why people should care.
}%note
Efficient storage is a core challenge faced by today's computational systems, especially due to the fast-growing high resolution acquisition/display devices, the demand for 3D data quality, and machine learning.
Thus, streaming services and remote cloud-based rendering (such as with Google Stadia and Nvidia Shield) have rapidly emerged in recent years. These storage-friendly systems have shown an unprecedented potential to reshape the traditional computer graphics pipeline by completely relieving the end user's computational workload and storage pressure, thus advancing device portability. Future spatial and immersive computing platforms such as VR and AR are potentially major beneficiaries due to the extensive demands of portability and energy-efficiency  \footnote{\url{https://www.vxchnge.com/blog/virtual-reality-data-storage-concern}}.

However, existing low-storage services have only been successfully deployed in relatively latency-insensitive platforms such as TV-based gaming consoles, due to the inevitable latency from transmission (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}). 
In comparison, VR and AR require remarkably low latency and significantly higher fidelity/speed for sickness-free and natural experience. As a consequence, there is a demand for massive high-quality spatial-temporal data. Unfortunately, no solution exists today to deliver high-quality and dynamic/interactive immersive content at low latency and with low local storage.

\note{
What prior works have done, and why they are not adequate.
(Note: this is just high level big ideas. Details should go to a previous work section.)
}%notez
%\qisun{(Jan 5, 2021) We should include two main fields: neural representation and foveated rendering}

Recent years have witnessed a surge in neural representations as an alternative to compact 3D representations such as point clouds and polygonal meshes, enabling compact assets. Application of neural representations include voxelization \cite{sitzmann2019deepvoxels}, distance fields \cite{sitzmann2019metasdf,mildenhall2020nerf,park2019deepsdf} and many more. However, existing solutions suffer from either high time consumption or low image fidelity with large scenes.

Acceleration without compromising perceptual quality is also an ultimate goal of real-time rendering.
Gaze-contingent rendering approaches have shown remarkable effectiveness of presenting imagery of perceptual quality identical to full resolution rendering~\cite{Guenter:2012:F3G,Patney:2016:TFR,Tursun:2019:LCA}. This is achieved by harnessing high field-of-view(FoV) head-mounted displays and high precision eye-tracking technologies. However, existing foveated rendering approaches still require locally stored full 3D assets, resulting in long transmission time and storage volume, whose availability is unfortunately limited in wearable devices.

\qisun{(Jan 22, 2021) Do we want to discuss 6DoF video synthesis here?}
\note{
What our method has to offer, sales pitch for concrete benefits, not technical details.
Imagine we are doing a TV advertisement here.
}%note

To overcome these open challenges mentioned above, we present the first gaze-contingent and interactive neural synthesis for high-quality 3D scenes. Aiming at the next-generation wearable, low-latency and low-storage VR/AR platforms, our compact neural model predicts stereo frames given users' real-time head and gaze motions. The predicted frames are perceptually identical to traditional real-time rendering of high-quality scenes.
With wide FoV and high resolution eye-tracked VR displays, our method significantly reduces 3D data storage (from 100MB to 0.5MB) thus enables instant viewing (without long loading/transmission waits) compared to traditional rendering pipelines; preserves visual quality, and improves responsiveness (from 9s to as low as 20 ms) compared with alternative view neural synthesis methods. 
Tailoring to future lightweight, power and memory-efficient wearable display systems, expected to be active all-day long akin to today's everyday-use mobile phones, our method achieves minimal local storage for high-quality 3D scenes.  %\praneeth{---- somoene please check the text in the brackets and make sure it is ok :)}

@@

To achieve immersive viewing with low storage, high visual quality as well as low systematic latency, we introduce spatial-angular human psychophysical characteristics to neural representation and synthesis. 
%To this end, we build both the retina-matched neural models and VR-tailored representation coordinates, towards high visual quality as well as low systematic latency. \praneeth{-- this line is a little confusing to me}. 
Specifically, we represent 3D scenes with concentric spherical coordinates. Our representation also allows for depicting the foveated color sensitivity and stereopsis, during both the training and inference phases. 
Furthermore, we derive an analytical spatial-temporal perception model to optimize our neural scene representation towards imperceptible loss in image quality and latency.\qisun{better? This paragraph was a bit messy before. Tried to clean up.}

@@

% In 3D scenes represented by concentric spherical coordinates, we model each camera ray's perceivablity on the retina. It particularly depicts the foveated color sensitivity and stereopsis during both training and inference. 
% By deriving an analytical spatial-temporal model, we further optimize our representation and networks towards imperceptible quality/precision loss and speed/latency.

%Specifically, we represent an arbitrary virtual 3D scene as a compact neural representation in a viewer centered inside-out fashion. 
%We account for the challenges such as low quality/resolution/performance common to existing 3D deep neural representation approaches, we further leverage the human visual perception to synthesize gaze-contingent and vision-matched immersive frames, without compromising the speed and image fidelity. 
%The method is further optimized with in the spatio-temporal domain to balance the latency and precision.

@@

We validate our approach by conducting numerical analysis and user evaluations on commercially available AR/VR display devices. The series of experiments reveal our method's effectiveness in delivering an interactive VR viewing experience, perceptually identical to an arbitrarily high quality rendered 3D scene, with significantly lowered data storage. We make the following major contributions:
\begin{itemize}
    \item A low-latency, interactive, and significantly low-storage immersive application. It also offers full perceptual quality with instant high resolution and high FoV first-person VR viewing.
    \item A scene parameterization and vision-matched neural representation/synthesis method considering both visual- and stereo- acuity.
    \item A spatiotemporally analytical model for jointly optimizing systematic latency and perceptual quality.
\end{itemize}

### 2021-01-28 TOG/2-prior.tex
last day

\subsection{Deep View Synthesis}

\subsection{Gaze-Contingent Rendering}
\cite{Guenter:2012:F3G,Patney:2016:TFR,Sun:2017:PGF,Kaplanyan:2019:DNR}

>>>>

\subsection{Image-based View Synthesis}
Image-based rendering (IBR) has been proposed in computer graphics to complement the traditional 3D assets and ray propagation pipelines \cite{levoy1996light,gortler1996lumigraph,shum2000review}. Benefiting from the uniform 2D representation, it delivers pixel-wise high quality without compromising performance. However, one major limitation of IBR is the demand for large and dense data acquisition, resulting in limited and sparse view ranges. 
To address this problem, recent years have witnessed extensive research on synthesizing novel image-based views instead of fully capturing or storing them. Examples include synthesizing light fields \cite{LearningViewSynthesis,Li2020LF,mildenhall2019llff} and multi-layer volumetric videos \cite{Broxton:immersiveLF}. However, despite being more compact, the synthesis is usually in a locally interpolated fashion than globally applicable with highly dynamic VR camera motions.

\subsection{Implicit Scene Representation for Neural Rendering}
To fully represent a 3D object and environment without large explicit storage, neural representations have drawn extensive attentions in both computer graphics and computer vision communities. 
With a deep neural network that depicts a 3D world as an implicit function, the neural networks may directly approximate an object's appearance given a camera pose \cite{sitzmann2019deepvoxels,sitzmann2019srns,sitzmann2019siren,mildenhall2020nerf}. 
Prior arts also investigated on implicitly representing shapes ~\cite{park2019deepsdf} and surfaces ~\cite{mescheder2019occupancy} to improve the visual quality of 3D objects. Inspired by \cite{park2019deepsdf}, signed distance function has been deployed for representing raw data such as point cloud ~\cite{atzmon2020sal,gropp2020implicit} with better efficiency ~\cite{chabra2020deep}. 
The input information was varied from 2D images ~\cite{lin2020sdf, yariv2020multiview}, 3D context ~\cite{saito2019pifu, oechsle2019texture}, time-varying vector field~\cite{niemeyer2019occupancy}, and local features~\cite{tretschk2020patchnets, liu2020neural}.
Current implicit representations primarily focus on locally ``out-side-in'' viewing of individual objects, with low field-of-view, speed and resolution. For large scenes, the lack of coverage may cause quality drop.
However, in virtual reality, the cameras are typically first-person and highly dynamic, requiring low latency, high resolution, and 6DoF coverage. 
%Our method, in comparison, is VR tailored via representing the scene with user-centered views.

\subsection{Gaze-Contingent Rendering}
Both rendering and neural synthesis suffer from the heavy computational load, thus systematic latency and lags.
With a full mesh representation, foveated rendering has been proposed to accelerate local rendering performance, including mesh- \cite{Guenter:2012:F3G,Patney:2016:TFR} or image-based \cite{Sun:2017:PGF,Kaplanyan:2019:DNR} methods. The accelerations are typically achieved through high field-of-view displays (such as VR head-mounted displays) and eye-tracking technologies. However, these all require full access to the original 3D assets or per-frame images. Consequently, the systems usually require large local storage, long transmission waits, or constant frame-based transmissions with latency. 
%On the other hand, most view synthesizing methods are designed for offline viewing or processing, other than real-time or interactive consumption, especially with high resolution and FoV.  
Our method bridges human vision to deep synthesis. It accelerates neural synthesis beyond local rendering, enabling both computational and storage efficiency. It is suitable for future cloud-based virtual reality with instant full-quality viewing and interaction.
%\zh{seems not sufficient?}

\subsection{Panorama-Based 6DoF Immersive Viewing}
Panorama images and videos greatly fit VR applications thanks to their 360 FoV coverage. However, the main challenge is their single projection center in each frame, limiting free-form camera translation, thus 6DoF immersive viewing experience. 
Recently, extensive research has been proposed to address this problem. Serrano et al. \shortcite{Serrano_TVCG_VR-6dof} presented a depth-based dis-occlusion method that dynamically reprojects to 6DoF cameras, enabling natural viewing with motion parallax. Pozo et al. \shortcite{Pozo:2019:I6V} jointly optimizes capture and viewing processes.
Machine learning approaches have advanced the robustness of various geometric and lighting conditions \cite{Attal:2020:ECCV,Lin:DeepPanorama,Benjamin:2020:RTV}. 
As a spherical image, panoramas are designed for capturing physical worlds with flexible reprojections to displays. The 6DoF viewing typically suffer from trade-offs among performance, achievable resolution, and dis-occlusion effects due to lack of knowledge of the 3D space.
With an orthogonal mission, our model represents and reproduces arbitrary 3D virtual worlds for low storage VR. It lowers memory usage by replacing traditional mesh- or volume-based data than predicting image-space occlusions.
